Arrows for Parallel Computation

### 2018-07-26 paper/content/topologySkeletons.tex

The parallel |pipe| skeleton is semantically equivalent to folding over a list |[arr a a]| of Arrows with |>>>|, but does this in parallel, meaning that the Arrows do not have to reside on the same thread/machine. We implement this skeleton using the |ArrowLoop| type class which gives us the |loop :: arr (a, b) (c, b) -> arr a c| combinator which allows us to express recursive fix-point computations in which output values are fed back as input. For example %this
\begin{code}
loop (arr (\(a, b) -> (b, a:b)))
\end{code}
which is the same as
\begin{code}
loop (arr snd &&& arr (uncurry (:)))
\end{code}
defines an Arrow that takes its input |a| and converts it into an infinite stream |[a]| of it. Using |loop| to our advantage gives us a first draft of a pipe implementation (Fig.~\ref{fig:pipeSimple}) by plugging in the parallel evaluation call |evalN conf fs| inside the second argument of |&&&| and then only picking the first element of the resulting list with |arr last|.

\begin{figure}[t]
\begin{code}
pipeSimple :: (ArrowLoop arr, ArrowLoopParallel arr a a conf) =>
	conf -> [arr a a] -> arr a a
pipeSimple conf fs =
    loop (arr snd &&&
        (arr (uncurry (:) >>> lazy) >>> loopParEvalN conf fs)) >>>
    arr last
\end{code}
\caption{Simple |pipe| skeleton. The use of |lazy| (Fig.~\ref{fig:edenlazyrightrotate}) is essential as without it programs using this definition would never halt. We need to enforce that the evaluation of the input |[a]| terminates before passing it into |evalN|.}
\label{fig:pipeSimple}
\end{figure}

>>>>

The parallel |pipe| skeleton is semantically equivalent to folding over a list |[arr a a]| of Arrows with |>>>|, but does this in parallel, meaning that the Arrows do not have to reside on the same thread/machine. We implement this skeleton using the |ArrowLoop| type class which gives us the |loop :: arr (a, b) (c, b) -> arr a c| combinator which allows us to express recursive fix-point computations in which output values are fed back as input. For example %this
\begin{code}
loop (arr (\(a, b) -> (b, a:b)))
\end{code}
which is the same as
\begin{code}
loop (arr snd &&& arr (uncurry (:)))
\end{code}
defines an Arrow that takes its input |a| and converts it into an infinite stream |[a]| of it. Using |loop| to our advantage gives us a first draft of a pipe implementation (Fig.~\ref{fig:pipeSimple}) by plugging in the parallel evaluation call |loopParEvalN conf fs| inside the second argument of |&&&| and then only picking the first element of the resulting list with |arr last|.

\begin{figure}[t]
\begin{code}
pipeSimple :: (ArrowLoop arr, ArrowLoopParallel arr a a conf) =>
	conf -> [arr a a] -> arr a a
pipeSimple conf fs =
    loop (arr snd &&&
        (arr (uncurry (:) >>> lazy) >>> loopParEvalN conf fs)) >>>
    arr last
\end{code}
\caption{Simple |pipe| skeleton. The use of |lazy| (Fig.~\ref{fig:edenlazyrightrotate}) is essential as without it programs using this definition would never halt. We need to ensure that the evaluation of the input |[a]| is not forced fully before passing it into |loopParEvalN|.}
\label{fig:pipeSimple}
\end{figure}
